---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Loading and preprocessing the data
First we should download the zipfile using the URL provided and read the CSV file after unzipping it.

```{r}
ActivityURL<- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(ActivityURL, destfile = "./activity.zip", method = "curl")
unzip("activity.zip")
activityData<- read.csv("activity.csv")
```

The date coloumn is currently classed a factor, we can change it to the date class

```{r}
activityData$date<- as.Date(activityData$date,format= "%Y-%m-%d")
```


## What is mean total number of steps taken per day?

We can find out the total number of steps taken per a day and present them in a histogram, using the ggplot2 and dplyr packages.

```{r histogram}
library(ggplot2)
library(dplyr)
dailySteps<- activityData %>% group_by(date) %>% summarise(total= sum(steps, na.rm = TRUE))

ggplot(dailySteps, aes(date, total, group= date))+
  geom_col()+
  labs(title = "Total Steps per a Day", x= "Date", y= "Total Steps")
```

We can then calculate the average total steps per day

```{r}
mean(dailySteps$total)
```

and the median total number of steps a day

```{r}
median(dailySteps$total)
```

## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
