---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Loading and preprocessing the data
First we should download the zipfile using the URL provided and read the CSV file after unzipping it.

```{r cache=TRUE}
ActivityURL<- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(ActivityURL, destfile = "./activity.zip", method = "curl")
unzip("activity.zip")
activityData<- read.csv("activity.csv")
```

The date coloumn is currently classed a factor, we can change it to the date class

```{r}
activityData$date<- as.Date(activityData$date,format= "%Y-%m-%d")
```


## What is mean total number of steps taken per day?

We can find out the total number of steps taken per a day and present them in a **histogram**, using the ggplot2 and dplyr packages.

```{r warning=FALSE, message=FALSE}
library(ggplot2)
library(dplyr)
```

```{r histogram}
dailySteps<- activityData %>% group_by(date) %>% summarise(total= sum(steps, na.rm = TRUE))

ggplot(dailySteps, aes(date, total, group= date))+
  geom_col()+
  labs(title = "Total Steps per a Day", x= "Date", y= "Total Steps")
```

We can then calculate the **average** total steps per day

```{r}
mean(dailySteps$total)
```

and the **median** total number of steps a day

```{r}
median(dailySteps$total)
```

## What is the average daily activity pattern?

We can plot the **average** number of steps for each 5 minute interval in a day by making a **time series plot**

```{r time series plot}
intervalSteps<- activityData %>% group_by(interval) %>% summarise(average= mean(steps, na.rm = TRUE))
with(intervalSteps, plot(interval, average, type = "l", xlab = "5 Minute Interval", ylab = "Average number of steps"))
```

And with this data we can find out which 5 minute interval has the **highest average** for a day


```{r}
intervalSteps$interval[which.max(intervalSteps$average)]
```


## Imputing missing values

There are many  missing values in the data, we can find out the exact number to see how much of an effect they have on the analysis

``` {r}
table(is.na(activityData$steps))
```

As we can see there is a significant amount, so we can try and replace the missing values with a logical alternative, such as the average for each interval which we calculated previously. We can make a new **histogram** with the NA values replaced by the **mean number of steps of the respective interval**

```{r histogram2}
activityData[,1][is.na(activityData[,1])]<- intervalSteps$average
dailyStepsWithmean<- activityData %>% group_by(date) %>% summarise(total= sum(steps))
ggplot(dailyStepsWithmean, aes(date, total))+
   geom_col()+
   labs(title = "Total Steps per a Day", x= "Date", y= "Total Steps")
```

It is possible to see in the histogram that there is less gaps when we replace the NA values, we can also test if the **mean** and **median** are affected

```{r mean and median}
mean(dailyStepsWithmean$total)
median(dailyStepsWithmean$total)
```

We now see that the mean and median have converged, which is not suprising as we replaced a significant amount of missing values with their corresponding average value
## Are there differences in activity patterns between weekdays and weekends?


